Solution to ``Chapter 2: Basic tail and concentration bounds"

### 2020-07-23 skills/math/tail_bound/tailbound_solution.tex
upload

\usepackage{comment}
% Definitions of handy macros can go here
\usepackage{amsmath,amssymb,amsthm,bm,mathtools}
\usepackage{multirow}
\usepackage{extarrows}
\usepackage{dsfont,multirow,hyperref,setspace,natbib,enumerate}
%\usepackage{dsfont,multirow,hyperref,setspace,enumerate}
\hypersetup{colorlinks,linkcolor={blue},citecolor={blue},urlcolor={red}} 
\usepackage{algpseudocode,algorithm}
\algnewcommand\algorithmicinput{\textbf{Input:}}
\algnewcommand\algorithmicoutput{\textbf{Output:}}
\algnewcommand\INPUT{\item[\algorithmicinput]}
\algnewcommand\OUTPUT{\item[\algorithmicoutput]}

@@

@@

\section{Summary}
\begin{thm}[Markov's inequality]\label{thm:markov}
	Suppose $X \geq 0$ is a random variable  with finite mean, we have
	\begin{equation}\label{eq:thmmarkov}
		\bbP(X\geq t) \leq \frac{E[X]}{t},\quad \text{ for all } \ t > 0.
	\end{equation}  
\end{thm}

\begin{thm}[Chebyshev's inequality]\label{thm:cheb}
	Suppose $X \geq 0$ is a random variable  with finite mean $\mu$ and finite variance, we have
	\begin{equation}\label{eq:cheby}
		\bbP(|X - \mu |\geq t) \leq \frac{var(X)}{t^2},\quad \text{ for all } \ t > 0.
	\end{equation}  
\end{thm}

\begin{thm}[Markov's inequality for polynomial moments]\label{thm:polymoment}
	Suppose the random variable $X$ has a central moment of order $k$. Applying Markov's inequality to the random variable $|X - \mu|^k$ yields
	\begin{equation}\label{eq:poly}
		\bbP(|X - \mu |\geq t) \leq \frac{\bbE[|X-\mu|^k]}{t^k},\quad \text{ for all } \ t > 0.
	\end{equation}  
\end{thm}

\begin{thm}[Chernoff bound]\label{thm:chernoff}
	Suppose the random variable $X$ has a moment generating function in the neighborhood of 0; i.e.\, $\varphi_X(\lambda) = \bbE[e^{\lambda X}] < +\infty,$ for all $\lambda \in (-b,b)$ with some $b>0$. Applying Markov's inequality to the random variable $Y = e^{\lambda(X- \mu)}$ yields
	\begin{equation}\label{eq:chernoff1}
		\bbP((X - \mu) \geq t) \leq \frac{\bbE[e^{\lambda (X-\mu)}]}{e^{\lambda t}}, \quad \text{ for all } \lambda\in (-b,b).
	\end{equation}
	Optimizing the choice of $\lambda$ for the tightest bound yields the Chernoff bound
	\begin{equation}\label{eq:chernoff2}
		\bbP((X - \mu) \geq t) \leq \inf_{\lambda \in [0,b)} \frac{\bbE[e^{\lambda (X-\mu)}]}{e^{\lambda t}}.
	\end{equation}
\end{thm}

\begin{thm}[Hoeffding bound for bounded variable]\label{thm:boundvar}
Consider a random variable $X$ with mean $\mu = \bbE(X)$. Assume that $X$ is bounded and $X \in [a,b]$ almost surely, where $a, b$ are two constants. Then, for any $\lambda \in \bbR$, we have
\begin{equation}
	\bbE[e^{\lambda X}] \leq e^{\frac{s(b-a)^2}{8}}.
\end{equation}
Particularly, the variable $X \sim subG(\frac{(b-a)^2}{4})$.
\end{thm}

@@

\begin{thm}[Moment of sub-Gaussian variable]\label{thm:momsubg}
Let $X \sim sG(\sigma^2)$. For all integer $k \geq 1$, we have
\begin{equation}\label{eq:thmmomsubg}
	\bbE[|X|^k] \leq  k 2^{k/2} \sigma^k \Gamma(\frac{k}{2} ),
\end{equation} 
where the Gamma function is defined as $\Gamma(x) = \int_{0}^{+\infty} t^{x -1} e^{-t} dt$. 
\end{thm}

\begin{thm}[One-sided Bernstein’s inequality]\label{thm:oneber}
 Let $X$ be a random variable. If $X \leq b$ almost surely, then 
\begin{equation}\label{eq:thmoneber1}
	\bbE[e^{\lambda (X-\bbE[X])}] \leq exp\left\{ \frac{\lambda^2 \bbE[X^2]/2}{1 - b\lambda /3}  \right\}, \quad \text{for all } \lambda \in [0, 3/b).
\end{equation}
Consequently, suppose there are $n$ independent variables $X_i \leq b$ almost surely, for all $i \in [n]$. We have
\begin{equation}\label{eq:thmoneber2}
	\bbP[\sum_{i = 1}^n (X_i - \bbE[X_i] ) \geq n\delta ] \leq exp\left\{ - \frac{n\delta^2}{\sum_{i = 1}^n \bbE[X_i^2] /n + b \delta /3} \right\},\quad \text{ for all } \delta \geq 0.
\end{equation}  
Particularly, suppose there are $n$ independent non-negative variables $Y_i \geq 0$, for all $i \in [n]$. The equation~\eqref{eq:thmoneber2} becomes
\begin{equation}\label{eq:thmonber0}
	\bbP[\sum_{i = 1}^n (Y_i - \bbE[Y_i] ) \leq n\delta ] \leq exp\left\{ - \frac{n\delta^2}{\sum_{i = 1}^n \bbE[Y_i^2] /n} \right\},\quad \text{ for all } \delta \geq 0.
\end{equation} 
\end{thm}

@@

\subsection{Exercise 2.1} (Tightness of inequalities.) The Markov's and Chebyshev's inequalities can not be improved in general. 
\begin{enumerate}
	\item[(a)] Provide a random variable $X \geq 0$ that attains the equality in Markov's inequality~\eqref{eq:thmmarkov}.  
	\item[(b)] Provide a random variable $Y$ that attains the equality in Chebyshev's inequality~\eqref{eq:cheby}.  
\end{enumerate}

@@

\item[(a)] For a given constant $t >0$, we define a variable $Y_{t} = X - t\mI_{[X\geq t]}$, where $\mI$ is the indicator function. Noted that $Y_{t}$ is a nonnegative variable, the Markov's inequality follows by taking expectation to $Y$,
\begin{equation}
	\bbE[Y_t] = \bbE[X] - t \bbP[X \geq t] \geq 0.
\end{equation}
Therefore, Markov's inequality meets the equality if and only if the expectation $\bbE[Y_t] = 0$. Since $Y_t$ is nonnegative, we have $\bbP(Y_t = 0) = 1$. Note that $Y_t = 0$ if and only if $X = 0$ or $X =t$.

@@

\item[(b)]  Chebyshev's inequality follows by applying Markov's inequality to the non-negative random variable $Z = {(X - \bbE[X])^2}$. Similarly as part (a), given a constant $t > 0$, the variable $Z = {(X - \bbE[X])^2}$ with distribution $\bbP(Z \in \{ 0,t^2 \}) = 1$ attains the equality of the Markov's inequality for $Z$ and the Chebyshev's inequality for $X$. By transformation, the distribution of $X$ satisfies the followings formula,
\begin{align}
\bbP(X = x) = 
	\begin{cases}
		p & \text{if } x = c,\\
		\frac{1-p}{2} & \text{if } x = c-t \text{ or } x = c+t, \\
		0 & \text{ otherwise },
	\end{cases}
	\end{align}
where $c \in \bbR$ is a constant and $p \in [0,1]$.

@@

\begin{rmk}[Tightness of Markov's inequality]
	Only a few variables can attain the equalities in Markov's and Chebyshev's inequalities. In research, we should pay attention to the concentration bounds tighter than Markov's inequality.
\end{rmk}

\subsection{Exercise 2.2}
\begin{lem}[Standard normal distribution]\label{lem:n01}
	Let $\phi(z)$ be the density function of a standard normal variable $Z \sim N(0,1)$. Then, 
	\begin{equation}\label{eq:e22a}
		\phi'(z) + z \phi(z) = 0,
	\end{equation}
	and \begin{equation}\label{eq:e22b}
		\phi(z)(\frac{1}{z} - \frac{1}{z^3} ) \leq \bbP(Z \geq z) \leq \phi(z)(\frac{1}{z} - \frac{1}{z^3}+ \frac{3}{z^5} ), \quad \text{for all }z>0.
	\end{equation}
\end{lem}
\begin{proof}First, we prove the equation~\eqref{eq:e22a}.

The pdf of standard normal distribution is
\[  \phi(z) = \frac{1}{\sqrt{2\pi}} \exp(-\frac{z^2}{2}  ).\]

The equation~\eqref{eq:e22a} follows by taking the derivative of $\phi(z)$.
\[\quad \phi'(z) = -z \frac{1}{\sqrt{2\pi}} \exp(-\frac{z^2}{2} ) = -z \phi(z). \] 
	Next, we prove the equation~\eqref{eq:e22b}.

	We write the upper tail probability of the standard normal variable as
	\begin{equation}\label{eq:e22pro1}
		\bbP(Z \geq z) = \int_{z}^{+\infty} \phi(t) dt = \int_{z}^{+\infty} - \frac{1}{t} \phi'(t) dt = \frac{1}{z} \phi(z) - \int_{z}^{+\infty} \frac{1}{t^2} \phi(t) dt,  
	\end{equation}
	where the second equality follows by the equation~\eqref{eq:e22b}. Applying the equation~\eqref{eq:e22b} to the last term in equation~\eqref{eq:e22pro1} yields
	\begin{equation}\label{eq:e22pro2}
	\int_{z}^{+\infty} \frac{1}{t^2} \phi(t) dt = \int_{z}^{+\infty} \frac{1}{t^3} \phi'(t) dt = - \frac{1}{z^3} \phi(z) + \int_{z}^{+\infty} \frac{3}{t^4} \phi(t) dt \geq  - \frac{1}{z^3} \phi(z)
	\end{equation}
	Plugging the equation~\eqref{eq:e22pro2} into the equation~\eqref{eq:e22pro1} , we have $\bbP(Z \geq z) \geq \phi(z)(\frac{1}{z}  - \frac{1}{z^3} )$. On the other hand, applying the equation~\eqref{eq:e22b} again to the equation~\eqref{eq:e22pro2} yields
	\begin{equation}\label{eq:e22pro3}
		\int_{z}^{+\infty} \frac{3}{t^4} \phi(t) dt = \int_{z}^{+\infty} - \frac{3}{t^5} \phi'(t) dt =\frac{3}{ z^5} \phi(z) -  \int_{z}^{+\infty} \frac{15}{t^6} \phi(t) dt \leq \frac{3}{ z^5} \phi(z).
	\end{equation}

	Combing equations~\eqref{eq:e22pro1}, \eqref{eq:e22pro2}, and  \eqref{eq:e22pro3}, we have $\bbP(Z \geq z) \leq \phi(z)(\frac{1}{z}  - \frac{1}{z^3} + \frac{3}{z^5} ).$
\end{proof}

@@

\begin{lem}[Polynomial bound and Chernoff bound]\label{lem:momcher}
	Let $X \geq 0$ be a non-negative variable. Suppose that the moment generating function of $X$, denoted $\varphi_{X}(\lambda)$, exists in the neighborhood of $\lambda = 0$. Given some $\delta > 0$, we have
	\begin{equation}\label{eq:lempoly}
		\inf_{k \in \mathbb{Z_+} } \frac{\bbE[|X|^k]}{\delta^k} \leq \inf_{\lambda >0} \frac{\bbE[e^{\lambda X}]}{e^{\lambda \delta}}. 
	\end{equation}

@@

\begin{proof}
	By power series, we have
	\begin{equation}\label{eq:power}
		 e^{\lambda X} = \sum_{k = 0}^{+\infty } \frac{X^k \lambda^k }{k!}, \quad \text{ for all } \lambda \in \bbR.
	\end{equation}
	Since the moment generating function $\varphi_X(\lambda)$ exists in the neighborhood of $\lambda = 0$, there exists a constant $b > 0$ such that 
		\[   \bbE [e^{\lambda X} ]= \sum_{k = 0}^{+\infty } \frac{ \bbE[|X|^k]  \lambda^k }{k!} < +\infty,\quad \text{ for all } \lambda \in (0,b). \]
	Therefore, the moment  $\bbE[|X|^k]$ exists for all $ k \in \mathbb{Z_+}$. Applying power series~\eqref{eq:power} to the right hand side of equation~\eqref{eq:lempoly} yields
	\begin{equation}\label{eq:powertopoly}
		\inf_{\lambda >0} \frac{\bbE[e^{\lambda X}]}{e^{\lambda \delta}} = \frac{ \sum_{k = 0}^{+\infty } \frac{ \bbE[|X|^k]  \lambda^k }{k!} }{ \sum_{k = 0}^{+\infty } \frac{\lambda^k \delta^k }{k!}  }.	\end{equation}
	By Cauchy's third inequality, we have
	\begin{equation}\label{eq:cauchy3}
		\frac{ \sum_{k = 0}^{+\infty } \frac{ \bbE[|X|^k]  \lambda^k }{k!} }{ \sum_{k = 0}^{+\infty } \frac{\lambda^k \delta^k }{k!}  } \geq \inf_{k \in \mathbb{Z}_{+} } \frac{\bbE[|X|^k]}{\delta^k}  
	\end{equation}
	Therefore, the equation~\eqref{eq:lempoly} follows by combining the equation~\eqref{eq:powertopoly} with equation~\eqref{eq:cauchy3}.	
\end{proof}

\begin{rmk}
	Applying different functions $g(X)$ to the Markov's inequality leads to different bounds for the tail probability of variable $X$. Equation~\eqref{eq:lempoly} implies that optimized polynomial bound is tighter than the Chernoff bound, provided that the moment generating function of $X$ exsits.  
\end{rmk}

\subsection{Exercise 2.4}\label{pro:boundvar}
In Exercise~\ref{pro:boundvar}, we prove the Theorem~\ref{thm:boundvar},  the Hoeffding bound for a bounded variable.

@@

Suppose $X$ is a bounded random variable, where $X \in [a,b]$ almost surely, and $a 
\leq b \in \bbR$ are two constants. Let $\mu = \bbE[X]$. Define the function 
\[ g(\lambda) = \log \bbE[e^{\lambda X }], \quad  \text{ for all } \lambda \in \bbR. \]
Applying Taylor Expansion to $g(\lambda)$ at 0, we have
\begin{equation}\label{eq:taylor}
	g(\lambda) =  g(0) + g'(0) \lambda  + \frac{g''(\lambda_0) }{2} \lambda^2, \quad \text{where } \lambda_0 = t \lambda, \text{ for some } t \in [0,1]. 
\end{equation}

>>>>

\usepackage{comment}
% Definitions of handy macros can go here
\usepackage{amsmath,amssymb,amsthm,bm,mathtools}
\usepackage{multirow}
\usepackage{extarrows}
\usepackage{dsfont,multirow,hyperref,setspace,natbib,enumerate}
%\usepackage{dsfont,multirow,hyperref,setspace,enumerate}
\hypersetup{colorlinks,linkcolor={blue},citecolor={blue},urlcolor={red}} 
\usepackage{algpseudocode,algorithm}
\algnewcommand\algorithmicinput{\textbf{Input:}}
\algnewcommand\algorithmicoutput{\textbf{Output:}}
\algnewcommand\INPUT{\item[\algorithmicinput]}
\algnewcommand\OUTPUT{\item[\algorithmicoutput]}
\usepackage{dsfont}
\usepackage{colonequals}

@@

\DeclareMathOperator{\var}{var}
\DeclareMathOperator{\subG}{subG}

@@

\section{Summary}
\begin{thm}[Markov's inequality]\label{thm:markov}
	Let $X \geq 0$ be a random variable  with finite mean. We have
	\begin{equation}\label{eq:thmmarkov}
		\bbP(X\geq t) \leq \frac{\bbE[X]}{t},\quad \text{ for all } \ t > 0.
	\end{equation}  
\end{thm}

\begin{thm}[Chebyshev's inequality]\label{thm:cheb}
	Let $X \geq 0$ be a random variable  with finite mean $\mu$ and finite variance. We have
	\begin{equation}\label{eq:cheby}
		\bbP(|X - \mu |\geq t) \leq \frac{\var(X)}{t^2},\quad \text{ for all } \ t > 0.
	\end{equation}  
\end{thm}

\begin{thm}[Markov's inequality for polynomial moments]\label{thm:polymoment}
	Let $X$ be a random variable. Suppose that the order $k$ central moment of $X$ exists. Applying Markov's inequality to the random variable $|X - \mu|^k$ yields
	\begin{equation}\label{eq:poly}
		\bbP(|X - \mu |\geq t) \leq \frac{\bbE\left[|X-\mu|^k\right]}{t^k},\quad \text{ for all } \ t > 0.
	\end{equation}  
\end{thm}

\begin{thm}[Chernoff bound]\label{thm:chernoff}
 Let $X$ be a random variable. Suppose that the moment generating function of $X$, denoted $\varphi_X(\lambda)$, exists in the neighborhood of 0; i.e., $\varphi_X(\lambda) = \bbE[e^{\lambda X}] < +\infty,$ for all $\lambda \in (-b,b)$ with some $b>0$. Applying Markov's inequality to the random variable $Y = e^{\lambda(X- \mu)}$ yields
	\begin{equation}\label{eq:chernoff1}
		\bbP((X - \mu) \geq t) \leq \frac{\bbE\left[e^{\lambda (X-\mu)}\right]}{e^{\lambda t}}, \quad \text{ for all } \lambda\in (-b,b).
	\end{equation}
	Optimizing the choice of $\lambda$ for the tightest bound, we obtain the Chernoff bound
	\begin{equation}\label{eq:chernoff2}
		\bbP((X - \mu) \geq t) \leq \inf_{\lambda \in [0,b)} \frac{\bbE\left[e^{\lambda (X-\mu)}\right]}{e^{\lambda t}}.
	\end{equation}
\end{thm}

\begin{thm}[Hoeffding bound for bounded variable]\label{thm:boundvar}
Let $X$ be a random variable with mean $\mu = \bbE(X)$. Suppose that $X \in [a,b]$ almost surely, where $a\leq b\in\bbR$ are two constants. Then, we have
\begin{equation}
	\bbE[e^{\lambda X}] \leq e^{\frac{s(b-a)^2}{8}}, \quad \text{for all } \lambda \in \bbR.
\end{equation}
Consequently, the variable $X \sim \subG\left(\frac{(b-a)^2}{4}\right)$.
\end{thm}

@@

\begin{thm}[Moment of sub-Gaussian variable]\label{thm:momsubg}
Let $X \sim \subG(\sigma^2)$. For all integer $k \geq 1$, we have
\begin{equation}\label{eq:thmmomsubg}
	\bbE[|X|^k] \leq  k 2^{k/2} \sigma^k \Gamma(\frac{k}{2} ),
\end{equation} 
where the Gamma function is defined as $\Gamma(x) = \int_{0}^{+\infty} t^{x -1} e^{-t} dt$. 
\end{thm}

\begin{thm}[One-sided Bernstein’s inequality]\label{thm:oneber}
 Let $X$ be a random variable. Suppose $X \leq b$ almost surely. We have 
\begin{equation}\label{eq:thmoneber1}
	\bbE\left[e^{\lambda (X-\bbE[X])}\right] \leq exp\left\{ \frac{\lambda^2 \bbE[X^2]/2}{1 - b\lambda /3}  \right\}, \quad \text{for all } \lambda \in [0, 3/b).
\end{equation}
Consequently, let $X_i$ be independent variables, and $X_i \leq b$ almost surely, for all $i \in [n]$. We have
\begin{equation}\label{eq:thmoneber2}
	\bbP\left[\sum_{i = 1}^n (X_i - \bbE[X_i] ) \geq n\delta \right] \leq exp\left\{ - \frac{n\delta^2}{\sum_{i = 1}^n \bbE[X_i^2] /n + b \delta /3} \right\},\quad \text{ for all } \delta \geq 0.
\end{equation}  
Particularly, let $X_i$ be independent nonnegative variables, for all $i \in [n]$. The equation~\eqref{eq:thmoneber2} becomes
\begin{equation}\label{eq:thmonber0}
	\bbP\left[\sum_{i = 1}^n (Y_i - \bbE[Y_i] ) \leq n\delta \right] \leq exp\left\{ - \frac{n\delta^2}{\sum_{i = 1}^n \bbE[Y_i^2] /n} \right\},\quad \text{ for all } \delta \geq 0.
\end{equation} 
\end{thm}

\begin{defn}[Bernstein's condition]
Let $X$ be a random variable with mean $\mu = \bbE[X]$ and variance $ \sigma^2 = \var(X)$. We say $X$ satisfies the Bernstein's condition with parameter $b$ if
\begin{equation}
	\left| \bbE [(X -\mu)^k ]\right| \leq \frac{1}{2} k! \sigma^2 b^{k-2}, \quad \text{for } k =3, 4,....
\end{equation} 
\end{defn}
Note that every bounded random variable $X$ satisfies the Bernstein's condition.

\begin{thm}[Bernstein-type bound] For any variable $X$ satisfying the Bernstein's condition, we have
\begin{equation}
	\bbE\left[e^{\lambda (X-\mu)}\right] \leq \exp\left\{ \frac{\lambda^2 \sigma^2}{2(1 - b |\lambda|)} \right\},\quad \text{for all } |\lambda| \leq \frac{1}{b}, 
\end{equation}
and the concentration inequality
\begin{equation}
	\bbP\left[|X - \mu| \geq t\right] \leq 2 \exp \left\{ -\frac{t^2}{2(\sigma^2 + bt)} \right\}, \quad \text{for all }t \geq 0.
\end{equation}

\end{thm}

@@

\subsection{Exercise 2.1} (Tightness of inequalities.) The Markov's and Chebyshev's inequalities are not able to be improved in general. 
\begin{enumerate}
	\item[(a)] Provide a random variable $X \geq 0$ that attains the equality in Markov's inequality~\eqref{eq:thmmarkov}.  
	\item[(b)] Provide a random variable $Y$ that attains the equality in Chebyshev's inequality~\eqref{eq:cheby}.  
\end{enumerate}

@@

\item[(a)] For a given constant $t >0$, we define a variable $Y_{t} = X - t \mathds{1}{[X\geq t]}$, where $\mathds{1}$ is the indicator function. Note that $Y_{t}$ is a nonnegative variable. The Markov's inequality follows by taking the expectation to $Y_t$,
\begin{equation}
	\bbE[Y_t] = \bbE[X] - t \bbP[X \geq t] \geq 0.
\end{equation}
Therefore, Markov's inequality meets the equality if and only if the expectation $\bbE[Y_t] = 0$. Since $Y_t$ is nonnegative, we have $\bbP(Y_t = 0) = 1$. Note that $Y_t = 0$ if and only if $X = 0$ or $X =t$.

@@

\item[(b)]  Chebyshev's inequality follows by applying Markov's inequality to the nonnegative random variable $Z = {(X - \bbE[X])^2}$. As in part (a), given a constant $t > 0$, the variable $Z = {(X - \bbE[X])^2}$ with distribution $\bbP(Z \in \{ 0,t^2 \}) = 1$ attains the equality of the Markov's inequality for $Z$. Consequently, the variable $X$ attains the equality of the Chebyshev's inequality for $X$. By transformation, the distribution of $X$ satisfies the followings formula,
\begin{align}
\bbP(X = x) = 
	\begin{cases}
		\quad p & \text{if } x = c,\\
		\quad \frac{1-p}{2} & \text{if } x = c-t \text{ or } x = c+t, \\
		\quad 0 & \text{ otherwise },
	\end{cases}
	\end{align}
where $c \in \bbR$ is a constant and $p \in [0,1]$.

@@

\begin{rmk}[Tightness of Markov's inequality]
	Only a few variables attain the equalities in Markov's and Chebyshev's inequalities. In research, we should pay attention to the concentration bounds tighter than Markov's inequality.
\end{rmk}

\subsection{Exercise 2.2}
\begin{lem}[Standard normal distribution]\label{lem:n01}
	Let $\phi(z)$ be the density function of a standard normal variable $Z \sim N(0,1)$. Then, 
	\begin{equation}\label{eq:e22a}
		\phi'(z) + z \phi(z) = 0,
	\end{equation}
	and \begin{equation}\label{eq:e22b}
		\phi(z)\left(\frac{1}{z} - \frac{1}{z^3} \right) \leq \bbP(Z \geq z) \leq \phi(z)\left(\frac{1}{z} - \frac{1}{z^3}+ \frac{3}{z^5} \right), \quad \text{for all }z>0.
	\end{equation}
\end{lem}
\begin{proof}First, we prove the equation~\eqref{eq:e22a}.

The pdf of the standard normal distribution is
\[  \phi(z) = \frac{1}{\sqrt{2\pi}} \exp\left(-\frac{z^2}{2}  \right).\]

The equation~\eqref{eq:e22a} follows by taking the derivative of $\phi(z)$. Specifically,
\[\quad \phi'(z) = -z \frac{1}{\sqrt{2\pi}} \exp\left(-\frac{z^2}{2} \right) = -z \phi(z). \] 
	Next, we prove the equation~\eqref{eq:e22b}.

	We write the upper tail probability of the standard normal variable as
	\begin{equation}\label{eq:e22pro1}
		\bbP(Z \geq z) = \int_{z}^{+\infty} \phi(t) dt = \int_{z}^{+\infty} - \frac{1}{t} \phi'(t) dt = \frac{1}{z} \phi(z) - \int_{z}^{+\infty} \frac{1}{t^2} \phi(t) dt,  
	\end{equation}
	where the second equality follows by the equation~\eqref{eq:e22a}. Applying the equation~\eqref{eq:e22a} to the last term in equation~\eqref{eq:e22pro1} yields
	\begin{equation}\label{eq:e22pro2}
	\int_{z}^{+\infty} \frac{1}{t^2} \phi(t) dt = \int_{z}^{+\infty} \frac{1}{t^3} \phi'(t) dt = - \frac{1}{z^3} \phi(z) + \int_{z}^{+\infty} \frac{3}{t^4} \phi(t) dt \geq  - \frac{1}{z^3} \phi(z)
	\end{equation}
	Plugging the equation~\eqref{eq:e22pro2} into the equation~\eqref{eq:e22pro1}, we obtain $\bbP(Z \geq z) \geq \phi(z)\left(\frac{1}{z}  - \frac{1}{z^3} \right)$. Applying the equation~\eqref{eq:e22a} again to the equation~\eqref{eq:e22pro2} yields
	\begin{equation}\label{eq:e22pro3}
		\int_{z}^{+\infty} \frac{3}{t^4} \phi(t) dt = \int_{z}^{+\infty} - \frac{3}{t^5} \phi'(t) dt =\frac{3}{ z^5} \phi(z) -  \int_{z}^{+\infty} \frac{15}{t^6} \phi(t) dt \leq \frac{3}{ z^5} \phi(z).
	\end{equation}

	Combing equations~\eqref{eq:e22pro1}, \eqref{eq:e22pro2} and  \eqref{eq:e22pro3}, we obtain $\bbP(Z \geq z) \leq \phi(z)\left(\frac{1}{z}  - \frac{1}{z^3} + \frac{3}{z^5} \right).$
\end{proof}

@@

\begin{lem}[Polynomial bound and Chernoff bound]\label{lem:momcher}
	Let $X \geq 0$ be a nonnegative variable. Suppose that the moment generating function of $X$, denoted $\varphi_{X}(\lambda)$, exists in the neighborhood of $\lambda = 0$. Given some $\delta > 0$, we have
	\begin{equation}\label{eq:lempoly}
		\inf_{k \in \mathbb{Z_+} } \frac{\bbE[|X|^k]}{\delta^k} \leq \inf_{\lambda >0} \frac{\bbE[e^{\lambda X}]}{e^{\lambda \delta}}. 
	\end{equation}

@@

\begin{proof}
	By power series, we have
	\begin{equation}\label{eq:power}
		 e^{\lambda X} = \sum_{k = 0}^{+\infty } \frac{X^k \lambda^k }{k!}, \quad \text{ for all } \lambda \in \bbR.
	\end{equation}
	Since the moment generating function $\varphi_X(\lambda)$ exists in the neighborhood of $\lambda = 0$, there exists a constant $b > 0$ such that 
		\[   \bbE [e^{\lambda X} ]= \sum_{k = 0}^{+\infty } \frac{ \bbE[|X|^k]  \lambda^k }{k!} < +\infty,\quad \text{ for all } \lambda \in (0,b). \]
	 Hence, the moment  $\bbE[|X|^k]$ exists, for all $ k \in \mathbb{Z_+}$. Applying power series~\eqref{eq:power} to the right hand side of equation~\eqref{eq:lempoly} yields
	\begin{equation}\label{eq:powertopoly}
		\inf_{\lambda >0} \frac{\bbE[e^{\lambda X}]}{e^{\lambda \delta}} = \frac{ \sum_{k = 0}^{+\infty } \frac{ \bbE[|X|^k]  \lambda^k }{k!} }{ \sum_{k = 0}^{+\infty } \frac{\lambda^k \delta^k }{k!}  }.	\end{equation}
	By Cauchy's third inequality, we have
	\begin{equation}\label{eq:cauchy3}
		\frac{ \sum_{k = 0}^{+\infty } \frac{ \bbE[|X|^k]  \lambda^k }{k!} }{ \sum_{k = 0}^{+\infty } \frac{\lambda^k \delta^k }{k!}  } \geq \inf_{k \in \mathbb{Z}_{+} } \frac{\bbE[|X|^k]}{\delta^k}  
	\end{equation}
	Therefore, we obtain the equation~\eqref{eq:lempoly} by combining the equation~\eqref{eq:powertopoly} with equation~\eqref{eq:cauchy3}.	
\end{proof}

\begin{rmk}
	Applying different functions $g(X)$ to the Markov's inequality leads to different bounds for the tail probability of variable $X$. Equation~\eqref{eq:lempoly} implies that optimized polynomial bound is at least as tight as the Chernoff bound, provided that the moment generating function of $X$ exsits.  
\end{rmk}

\subsection{Exercise 2.4}\label{pro:boundvar}
In Exercise~\ref{pro:boundvar}, we prove Theorem~\ref{thm:boundvar},  the Hoeffding bound for a bounded variable.

@@

Let $X$ be a bounded random variable, and $X \in [a,b]$ almost surely, where $a 
\leq b \in \bbR$ are two constants. Let $\mu = \bbE[X]$. Define the function 
\[ g(\lambda) = \log \bbE[e^{\lambda X }], \quad  \text{ for all } \lambda \in \bbR. \]
Applying Taylor Expansion to $g(\lambda)$ at 0, we have
\begin{equation}\label{eq:taylor}
	g(\lambda) =  g(0) + g'(0) \lambda  + \frac{g''(\lambda_0) }{2} \lambda^2, \quad \text{where } \lambda_0 = t \lambda, \text{ for some } t \in [0,1]. 
\end{equation}

In equation~\eqref{eq:taylor}, the term $g(0) =\log \bbE[e^{0}] = 0$. By power series \eqref{eq:power}, we obtain the first derivative $g'(\lambda)$ as follows,
\begin{align}
	g'(\lambda ) & = \left( \log \sum_{k = 0}^n \frac{\lambda^k }{k!} \bbE[X^k]  \right)' \\
	&= {\sum_{k = 0}^n \frac{ \lambda^k }{k!} \bbE[X^{(k+1)}] }\Bigg/{\sum_{k = 0}^n \frac{\lambda^k }{k!} \bbE[X^k]}  \label{eq:firstde} \\
	&= \frac{\bbE[X e^{\lambda X} ]}{\bbE[ e^{\lambda X} ]}.
\end{align}
Therefore, $g'(0) = \bbE[X] = \mu$. Taking the derivative to equation~\eqref{eq:firstde}, we obtain the second-order derivative $g''(\lambda)$ as follows,
\begin{align}
	g''(\lambda)
	&= {\sum_{k = 0}^n \frac{ \lambda^k }{k!} \bbE[X^{(k+2)}] }\Bigg/{\sum_{k = 0}^n \frac{\lambda^k }{k!} \bbE[X^k]} - \left( {\sum_{k = 0}^n \frac{ \lambda^k }{k!} \bbE[X^{(k+1)}] }\Bigg/{\sum_{k = 0}^n \frac{\lambda^k }{k!} \bbE[X^k]}  \right)^2\\
	&= \frac{\bbE[X^2 e^{\lambda X} ]}{\bbE[ e^{\lambda X} ]} - \left( \frac{\bbE[X e^{\lambda X} ]}{\bbE[ e^{\lambda X} ]} \right)^2.\\
	%& \overset{\Delta}{=} \bbE_{\lambda} [X^2] - (\bbE_{\lambda} [X])^2
\end{align}
We interpret the second-order derivative $g''(\lambda)$ as the variance of $X$ with the re-weighted distribution $d P' = {e^{\lambda X}}/{\bbE[ e^{\lambda X} ]} d P_X $, where $P_X$ is the distribution of $X$. Taking integral of 1 with respect to $dP'$, we have
\begin{equation}
	\int_{-\infty}^{+\infty} d P' = \int_{-\infty}^{+\infty}  \frac{e^{\lambda X}}{\bbE[ e^{\lambda X} ]} d P_X = 1,
\end{equation} 
which implies that the function $P'$ is a valid probability distribution. Under all possible re-weighted distributions, the variance of $X$ is upper bounded as follows,
\[ \var(X) = \var(X - \frac{a+b}{2}) \leq \bbE[(X - \frac{a+b}{2} )^2] \leq \frac{(b-a)^2}{4}, \]
where the term $\frac{(b-a)^2}{4}$ follows by letting $X$ supported on the boundaries $a$ and $b$ only. Hence, the second-order derivative $g''(\lambda) \leq \frac{(b-a)^2}{4}$. We plug the results of $g'$ and $g''$ into the equation~\eqref{eq:taylor}. Then,
\begin{equation}\label{eq:e24}
	g(\lambda) =  g(0) + g'(0) \lambda  + \frac{g''(\lambda_0) }{2} \lambda^2 \leq 0 + \lambda \mu+ \frac{(b-a)^2}{8} \lambda^2. 
\end{equation}
Taking the exponentiation on both sides of the inequality~\eqref{eq:e24}, we have
\begin{equation}\label{eq:boundvar}
	\bbE[e^{\lambda X}] = \exp(g( \lambda)) \leq e^{\mu \lambda + \frac{(b-a)^2}{8} \lambda^2}.
\end{equation}
The equation~\eqref{eq:boundvar} implies that $X$ is a sub-Gaussian variable with at most $\sigma = \frac{(b-a)}{2}$.
\end{proof}
